\LaTeXe\ guide for the Journal of Functional Programming

### 2019-04-26 papers/jfp/8-engineering.tex
Move self-tracking to S8.7 and add an implementation

% \vspace{1mm}
% \begin{minted}[xleftmargin=10pt]{haskell}
% data Key k     = Key k   | KeyTask k
% data Value v t = Value v | ValueTask t

% fetchValue :: Functor f => (Key k -> f (Value v t)) -> k -> f v
% fetchValue fetch key = extract <$> fetch (Key key)
%   where
%     extract (Value v) = v
%     extract _ = error "Inconsistent fetch"

% -- | Fetch a task description.
% fetchValueTask :: Functor f => (Key k -> f (Value v t)) -> k -> f t
% fetchValueTask fetch key = extract <$> fetch (KeyTask key)
%   where
%     extract (ValueTask t) = t
%     extract _ = error "Inconsistent fetch"

% -- | A model for 'Monad', works beautifully and allows storing the key on the
% -- disk.
% selfTrackingM :: forall k v t. (t -> Task Monad k v) -> Tasks Monad k t -> Tasks Monad (Key k) (Value v t)
% selfTrackingM _      _     (KeyTask _) = Nothing -- Task keys are inputs
% selfTrackingM parser tasks (Key     k) = runTask <$> tasks k
%   where
%     -- Fetch the task description, parse it, and then run the obtained task
%     runTask :: Task Monad k t -> Task Monad (Key k) (Value v t)
%     runTask act = Task $ \fetch -> do
%         task <- parser <$> run act (fetchValueTask fetch)
%         Value <$> run task (fetchValue fetch)
% \end{minted}
% \vspace{1mm}

@@

\noindent
Here we extend the usual mapping from file paths to file contents with an
additional key type \hs{GccVersion} mapped to a \hs{Version}. The task
abstraction needs to be adjusted to cope with such keys (the prefix \hs{T}
stands for ``typed''):

@@

newtype TTask c k v = TTask { run :: @\std{forall}@ f. c f => Fetch k f -> f v }

type TTasks c k = @\std{forall}@ v. k v -> Maybe (TTask c k v)
\end{minted}
\vspace{1mm}

\noindent
The changes compared to the definition in~\S\ref{sec-task} are minimal: (i) the
\hs{TTask} now uses a typed \hs{Fetch} callback (we define a separate type
synonym only for readability), and (ii) the type of \hs{TTasks} is now
polymorphic in \hs{v} instead of being parameterised by a concrete \hs{v}. The
example below demonstrates how \hs{fetch} can be used to retrieve dependencies
of different types: the rule \cmd{release.txt} concatenates the contents of two
\hs{File}s, while the rule \cmd{main.o} uses the numeric \hs{GccVersion} to
determine the way the \hs{source} file should be compiled.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
example :: TTasks Monad Key
example (File "release.txt") = Just $ TTask $ \fetch -> do
    readme  <- fetch (File "README")
    license <- fetch (File "LICENSE")
    return (readme ++ license)
example (File "main.o") = Just $ TTask $ \fetch -> do
    let source = "main.c"
    version <- fetch GccVersion
    if version >= Version 8 0 then compileNew source
                              else compileOld source
example _ = Nothing
\end{minted}
\vspace{1mm}

@@

>>>>

@@

\noindent
Here we extend the usual mapping from file paths to file contents with an
additional key type \hs{GccVersion} mapped to a \hs{Version}. The task
abstraction needs to be adjusted to cope with such keys (the suffix \hs{T}
stands for ``typed''):

@@

newtype TaskT c k v = TaskT { runT :: @\std{forall}@ f. c f => Fetch k f -> f v }

type TasksT c k = @\std{forall}@ v. k v -> Maybe (TaskT c k v)
\end{minted}
\vspace{1mm}

\noindent
The changes compared to the definition in~\S\ref{sec-task} are minimal: (i) the
\hs{TaskT} now uses a typed \hs{Fetch} callback (we define a separate type
synonym only for readability), and (ii) the type of \hs{TasksT} is now
polymorphic in \hs{v} instead of being parameterised by a concrete \hs{v}. The
example below demonstrates how \hs{fetch} can be used to retrieve dependencies
of different types: the rule \cmd{release.txt} concatenates the contents of two
\hs{File}s, while the rule \cmd{main.o} uses the numeric \hs{GccVersion} to
determine the way the \hs{source} file should be compiled.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
example :: TasksT Monad Key
example (File "release.txt") = Just $ TaskT $ \fetch -> do
    readme  <- fetch (File "README")
    license <- fetch (File "LICENSE")
    return (readme ++ license)
example (File "main.o") = Just $ TaskT $ \fetch -> do
    let source = "main.c"
    version <- fetch GccVersion
    if version >= Version 8 0 then compileNew source
                              else compileOld source
example _ = Nothing
\end{minted}
\vspace{1mm}

@@

\todo{AM/NM}{We are talking about sets above, but are using lists, which is a
bit inconsistent. We could switch from \hs{[k]} to \hs{Set k} fairly easily.
Shall we?}

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
data Key k v s a where
    Script :: k -> Key k v s s -- Keys for build scripts
    Value  :: k -> Key k v s v -- Keys for all other values

selfTracking :: (s -> Task Monad k v) -> Tasks Monad k s
                                      -> TasksT Monad (Key k v s)
selfTracking _     _     (Script _) = Nothing -- Scripts are inputs
selfTracking parse tasks (Value  k) = runScript <$> tasks k
  where
    -- Fetch the script, parse it, and then run the obtained task
    runScript :: Task Monad k s -> TaskT Monad (Key k v s) v
    runScript task = TaskT $ \fetch -> do
        script <- run task (fetch . Script)
        run (parse script) (fetch . Value)
\end{minted}
\vspace{1mm}
